Determinism, Complexity, and Predictability in Computer Performance

### 2014-07-16 model.tex
formatting errors

ARIMA forecasting is a common and time-tested procedure.  Its
adjustments for seasonality, nonstationarity, and noise make it an
appropriate choice for short-term predictions of time-series data
generated by a wide range of processes.  If information is being
generated and/or transmitted in a nonlinear way, however, a global
linear fit is inappropriate and ARIMA forecasts can be inaccurate.
Another weakness of this method is prediction horizon: an ARIMA
forecast is guaranteed to converge to the mean after some number of
predictions, depending on model order.  To sidestep this issue, we
build forecasts in a stepwise fashion: i.e., fit the \arima model
to the existing data, use that model to perform a one-step prediction,
rebuild it using the latest observations, and iterate until the
desired prediction horizon is reached.  (For consistency, we take the
same approach with the other three models in this study as well, even
though doing so amounts to artificially hobbling LMA.)

@@

\newcolumntype{C}{>{\centering\arraybackslash} m{0.6\columnwidth} }
\begin{figure*}
  \renewcommand{\arraystretch}{1.25}
  \begin{tabular}{m{0.1\columnwidth}|CCC}
    & {\large \col} & {\large \gcc} & {\large \svdfive} \\
    & & & \\
            \hline \\
    \rotatebox{90}{\large{random walk}} &
    \includegraphics[width=0.6\columnwidth]{figs/colRWForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccRWForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveRWForecast} \\
    \rotatebox{90}{\large \naive} &
    \includegraphics[width=0.6\columnwidth]{figs/colMeanForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccMeanForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveMeanForecast} \\
    \rotatebox{90}{\large \arima} &
    \includegraphics[width=0.6\columnwidth]{figs/colARIMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccARIMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveARIMAForecast} \\
    \rotatebox{90}{\large LMA} &
    \includegraphics[width=0.6\columnwidth]{figs/colLMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccLMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveLMAForecast}
  \label{fig:forecast-example}
  \end{tabular}
  \caption{Predicted ($p_j$) versus true values ($c_j$) for forecasts of \col,
     \gcc, and \svdfive and all four prediction strategies.
  }
  \label{fig:forecast-example}
\end{figure*}
In these images, the vertical axis is the prediction $p_j$ and the
horizontal axis is the true continuation $c_j$.  On such a plot, a
perfect prediction would lie on the diagonal.  LMA, for instance,
generates a very accurate prediction of the \col data, while \arima
does not.  Horizontal lines result when a constant predictor (e.g.,
\naive) is used on a non-constant signal.  Point clouds reflect the
structure of the distribution of the errors---roughly normally
distributed, for example, in the case of LMA on \gcc.  Note that the
shapes of some of the plots in Figure~\ref{fig:forecast-example}, random walk and \arima on \col for example,  are
reminiscent of the projected embedding in Figure~\ref{fig:embedding}a.
For a random-walk predictor, a $p_j$ vs. $c_j$ plot is equivalent to a
two-dimensional embedding with $\tau=1$.  For \arima, the
correspondence is not quite as simple, since the $p_j$ values are
linear combinations of a number of past values of the $c_j$, but the
effect is largely the same\footnote{The temporal \emph{ordering} of
  the points on the \arima $p_j$ vs. $c_j$ plot does not match that of
  a projected embedding of the time series, however.}.

>>>>

ARIMA forecasting is a common and time-tested procedure.  Its
adjustments for seasonality, nonstationarity, and noise make it an
appropriate choice for short-term predictions of time-series data
generated by a wide range of processes.  If information is being
generated and/or transmitted in a nonlinear way, however, a global
linear fit is inappropriate and ARIMA forecasts can be inaccurate.
Another weakness of this method is prediction horizon: an ARIMA
forecast is guaranteed to converge to the\alert{ mean, a constant, or a linear trend}, after some number of
predictions, depending on model order.  To sidestep this issue, we
build forecasts in a stepwise fashion: i.e., fit the \arima model
to the existing data, use that model to perform a one-step prediction,
rebuild it using the latest observations, and iterate until the
desired prediction horizon is reached.  (For consistency, we take the
same approach with the other three models in this study as well, even
though doing so amounts to artificially hobbling LMA.)

@@

\newcolumntype{C}{>{\centering\arraybackslash} m{0.6\columnwidth} }
\begin{figure*}
  \renewcommand{\arraystretch}{1.25}
  \begin{tabular}{m{0.1\columnwidth}|CCC}
    & {\large \col} & {\large \gcc} & {\large \svdfive} \\
    & & & \\
            \hline \\
    \rotatebox{90}{\large{random walk}} &
    \includegraphics[width=0.6\columnwidth]{figs/colRWForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccRWForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveRWForecast} \\
    \rotatebox{90}{\large \naive} &
    \includegraphics[width=0.6\columnwidth]{figs/colMeanForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccMeanForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveMeanForecast} \\
    \rotatebox{90}{\large \arima} &
    \includegraphics[width=0.6\columnwidth]{figs/colARIMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccARIMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveARIMAForecast} \\
    \rotatebox{90}{\large LMA} &
    \includegraphics[width=0.6\columnwidth]{figs/colLMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/gccLMAForecast} &
    \includegraphics[width=0.6\columnwidth]{figs/svdfiveLMAForecast}
  \end{tabular}
  \caption{Predicted ($p_j$) versus true values ($c_j$) for forecasts of \col,
     \gcc, and \svdfive and all four prediction strategies.
  }
  \label{fig:forecast-example}
\end{figure*}
In these images, the vertical axis is the prediction $p_j$ and the
horizontal axis is the true continuation $c_j$.  On such a plot, a
perfect prediction would lie on the diagonal.  LMA, for instance,
generates a very accurate prediction of the \col data, while \arima
does not.  Horizontal lines result when a constant predictor (e.g.,
\naive) is used on a non-constant signal.  Point clouds reflect the
structure of the distribution of the errors---roughly normally
distributed, for example, in the case of LMA on \gcc.  Note that the
shapes of some of the plots in Figure~\ref{fig:forecast-example}, random walk and \arima on \col for example,  are
reminiscent of the projected embedding in Figure~\ref{fig:embedding}a.
For a random-walk predictor, a $p_j$ vs. $c_j$ plot is equivalent to a
two-dimensional embedding with $\tau=1$.  For \arima, the
correspondence is not quite as simple, since the $p_j$ values are
linear combinations of a number of past values of the $c_j$, but the
effect is largely the same\footnote{The temporal \emph{ordering} of
  the points on the \arima $p_j$ vs. $c_j$ plot does not match that of
  a projected embedding of the time series, however.}.
